% !TEX encoding = UTF-8 Unicode
\documentclass[12pt]{amsart}

%\usepackage[notcite,notref]{showkeys}
\usepackage{tikz}
%\usepackage{bbm,bm}
%\usepackage{rotating}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{bm}
\usepackage[all]{xy}
%\usepackage{todonotes}



\sloppy
\flushbottom

\usepackage{color}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\green}[1]{\textcolor{green!75!black}{#1}}

\def\Pda{\Psi_{21}(\alpha)}

\usepackage[colorlinks=true,%linkcolor=RoyalBlue,
citecolor=cyan,urlcolor=blue,linktocpage=true,pagebackref]{hyperref} 


%%%%%%%%%%%%%%
\usepackage{xargs}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny,linecolor=red,backgroundcolor=red!25,bordercolor=red]{todonotes}
\setlength{\marginparwidth}{2cm}
\newcommandx{\nantel}[2][1=]{\todo[linecolor=brown,backgroundcolor=red!25,bordercolor=blue,#1]{#2 ---Nantel}}
\newcommandx{\jc}[2][1=]{\todo[linecolor=purple,backgroundcolor=blue!25,bordercolor=red,#1]{#2 ---JC}}
\newcommandx{\john}[2][1=]{\todo[linecolor=purple,backgroundcolor=green!25,bordercolor=red,#1]{#2 ---John}}


%%---- page geometry & headings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\addtolength{\textheight}{48pt}
\addtolength{\voffset}{-24pt}
\addtolength{\textwidth}{86pt}
\addtolength{\hoffset}{-42pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagestyle{headings} % empty}

\date{\today}

%-------------------------------------------------------------------
% ENVIRONMENTS
%-------------------------------------------------------------------
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}{\red{Conjecture}}
\newtheorem{property}{\red{Property}}

\theoremstyle{definition}
\newtheorem{example}[theorem]{Example} 
\newtheorem{remark}[theorem]{Remark} 
\newtheorem{definition}[theorem]{Definition} 

\theoremstyle{remark}
\newtheorem*{associativity}{Associativity}
%\newtheorem*{unitality}{Unitality}
%\newtheorem*{naturality}{Naturality}
\newtheorem*{compatibility}{Compatibility}
%\newtheorem*{connectedness}{Connectedness}
%\newtheorem*{multiplicativity}{Multiplicativity}
%\newtheorem*{warning}{Warning}

 
 
% TO SEE SPECIAL COMMENT "uncomment" the line bellow
% make sure that "\excludecomment{moredetails}" is the "commented" 
% TO HIDE SPECIAL COMMENT do the opposite as above
\specialcomment{moredetails}{\begingroup\color{red}}{\endgroup}
%\excludecomment{moredetails}


% SET: I,J,K,L
% Set composition:  A,B,C,D
% Parts O,P,Q,R,T,E,F

%-------------------------------------------------------------------
% MACROS
%-------------------------------------------------------------------
\newcommand{\draftnote}[1]{\marginpar{\raggedright\textsf{\hspace{0pt}\red{#1}}}}

\newcommand{\Vinc}{\begin{turn}{90}$\subseteq$\end{turn}} 
\newcommand{\Veq}{\begin{turn}{90}$=$\end{turn}}

\def\id{\mathrm{id}}
\def\Id{{\bf 1}}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\End}{End}

\def\llb{{[\![}}
\def\rrb{{]\!]}}


\newcommand{\abs}[1]{\lvert#1\rvert} %for absolute value
\newcommand{\pmat}[1]{\begin{pmatrix}#1\end{pmatrix}} %for round matrix

\newcommand{\larc}[1]{\hspace{-.4ex}\overset{#1}{\frown}\hspace{-.4ex}}
\newcommand{\slarc}[1]{\overset{#1}{\frown}}

\newcommand{\gel}{>_{\ell}}
\newcommand{\lel}{<_{\ell}}

\newcommand{\one}{\mathbf{1}}
\newcommand{\euler}{\mathbf{e}}

\let\onto=\twoheadrightarrow
\let\into=\hookrightarrow
\let\map=\xrightarrow
\newcommand{\xyinc}{\ar@{^{(}->}}
\newcommand{\mapsby}[1]{ \overset{#1}{\longmapsto}}

\newcommand{\type}[2]{\mathsf T_{#1}(#2)}

\def\field{\Bbbk}
\newcommand{\Sr}{\mathrm{S}} %symmetric group

\newcommand{\rH}{\mathrm{H}}
\newcommand{\rL}{{\mathbf l}}
\newcommand{\rG}{{\mathbf g}}
\newcommand{\rD}{\mathrm{D}}

\newcommand{\bL}{\mathbf L}
\newcommand{\bE}{\mathbf E}
\newcommand{\bH}{\mathbf H}
\newcommand{\bh}{\mathbf h}
\newcommand{\bk}{\mathbf k}
\newcommand{\bK}{\mathbf K}
\newcommand{\bPi}{\mathbf \Pi}
\newcommand{\bG}{\mathbf{G}} 
\newcommand{\bp}{\mathbf p}
\newcommand{\bq}{\mathbf q}
\newcommand{\bmm}{\mathbf m}%\bm exists
\newcommand{\bd}{\mathbf d}

\newcommand{\Fc}{\mathcal{F}}
\newcommand{\Tc}{\mathcal{T}}
\newcommand{\Kc}{\mathcal{K}}
\newcommand{\Kcb}{\overline{\Kc}}

\newcommand{\Nb}{\mathbb{N}}
\newcommand{\Qb}{\mathbb{Q}}
\newcommand{\Cb}{\mathbb{C}}
\newcommand{\Zb}{\mathbb{Z}}
\newcommand{\Fb}{\mathbb{F}}
\newcommand{\Fq}{\Fb_q}


\newcommand{\Qvar}{{\bm q}}
\newcommand{\Zvar}{{\bm z}}


\newcommand{\Nf}{\mathfrak{n}}
\newcommand{\bNf}{\overline{\Nf}}
\newcommand{\Mf}{\mathfrak{m}}
\newcommand{\bMf}{\overline{\Mf}}

\newcommand{\GL}{\mathrm{GL}}
\newcommand{\Un}{\mathrm{U}}
\newcommand{\Mat}{\mathrm{M}}
\newcommand{\Tri}{\mathrm{T}}

\newcommand{\FC}{\bm{\mathrm{f}}}
\newcommand{\CF}{\bm{\mathrm{cf}}}
\newcommand{\SC}{\bm{\mathrm{scf}}}

\newcommand{\bdot}{\bm\cdot}

\newcommand{\qand}{\quad\text{and}\quad}
\newcommand{\qor}{\quad\text{or}\quad}
\newcommand{\qqand}{\qquad\text{and}\qquad}
\newcommand{\qqor}{\qquad\text{or}\qquad}

%\newcommand{\jc}[1]{\todo[color=orange!30]{#1 \\ \hfill --- J-C.}}
%\newcommand{}[1]{\todo[color=red!30]{#1 \\ \hfill --- N.}}
\setlength{\marginparwidth}{2cm}

\newcommand{\PF}{\mathcal{PF}}
\newcommand\area{\mathop{\mathrm{area}}}
\newcommand{\dinv}{\mathop{\mathrm{dinv}}}
\newcommand{\ides}{\mathop{\mathrm{ides}}}
\newcommand{\des}{\mathop{\mathrm{des}}}
\newcommand{\desc}{\mathop{\mathrm{desc}}}
\newcommand{\ank}{\mathop{\mathrm{ank}}}
\newcommand{\ret}{\mathop{\mathrm{ret}}}
\newcommand{\comp}{\mathop{\mathrm{comp}}}
\newcommand{\ps}{\mathop{\mathrm ps^1}}

% math commands
\newcommand{\set}[2]{\left\{ #1 \;\middle|\; #2 \right\}} % set notation
\newcommand{\bigset}[2]{\big\{ #1 \;\big|\; #2 \big\}} % big set notation
\newcommand{\Bigset}[2]{\Big\{ #1 \;\Big|\; #2 \Big\}} % Big set notation
\newcommand{\setangle}[2]{\left\langle #1 \;\middle|\; #2 \right\rangle} % set notation
\newcommand{\ssm}{\smallsetminus} % small set minus
\newcommand{\dotprod}[2]{\left\langle \, #1 \; \middle| \; #2 \, \right\rangle} % dot product
\newcommand{\symdif}{\,\triangle\,} % symmetric difference
\newcommand{\eqdef}{\mbox{\,\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}}\ensuremath{=}\,}} % :=
\newcommand{\defeq}{\mbox{~\ensuremath{=}\raisebox{0.2ex}{\scriptsize\ensuremath{\mathrm:}} }} % =:
\newcommand{\simplex}{\triangle} % simplex
\renewcommand{\implies}{\Rightarrow} % imply sign
\newcommand{\transpose}[1]{{#1}^t} % transpose matrix

%%%%%%%%%%%%%%
%%  TITLE AND AUTHORS
%%%%%%%%%%%%%%
\author{Fran\c{c}ois Bergeron}

\author{Nantel Bergeron}\address[Bergeron]
{Department of Mathematics and Statistics\\ York  University\\ To\-ron\-to, Ontario M3J 1P3\\ CANADA}
\email{bergeron@mathstat.yorku.ca}
\urladdr{http://www.math.yorku.ca/bergeron}
\thanks{With partial support of Bergeron's York University Research Chair and NSERC}

\author{Cesar Caballos}
\author{Vincent Pilaud}

\title[Note and Conjectures on Multivariate Diagonal Harmonics]{Note and Conjectures on \\ Multivariate Diagonal Harmonics}

\begin{document}

\begin{abstract}
We try to summarize the advancement made in our work session in Paris, in February 2019.
\end{abstract}

\maketitle

\section{Paris, Feb 22, 2019}

\subsection{The square case~$n = m$}
We are studying the Frobenius image of the character of multivariate diagonal harmonics. Fran\c{c}ois shows that it is possible to find a formula
{\begin{equation}\label{eq:En}
{\mathcal E}_n(\Qvar;\Zvar) = \sum_{\lambda \vdash n \atop \nu \vdash \ell<{n\choose 2}} c_{\nu,\lambda} s_\nu(\Qvar) \otimes s_\lambda(\Zvar),
\end{equation}}
where the $c_{\mu,\lambda}$ are positive integers and the specialization $\Qvar \mapsto q_1 + q_2 + \dots + q_r$ gives the $r$-graded Frobenius image of the character of the multivariate diagonal harmonics with $r$ set of variables of cardinality $n$ with the diagonal action of~$S_n$.
This formula is computed explicitly by computer for~$n \in \{1,2,3,4,5,6\}$.

Guided by our heuristic that such formula such is constructed from the Tamari lattice using Hopf chain, we came to the conclusion that it is possible to write~${\mathcal E}_n(\Qvar;\Zvar)$ in different ways.
Let $\mu\subseteq\delta_n$ be a partition included in the staircase represent Dyck path in the Tamari lattice.
Possibly adding zero parts, we require that~$\ell(\mu) = n$ to encode~$n$.

\begin{conjecture}\label{conj:master}
Let $t$ be a variable, our multishuffle conjecture is
\begin{equation}\label{eq:multishuff}
\blue{
{\mathcal E}_n(\Qvar+t;\Zvar) = \sum_{\mu\subseteq\delta_n} \sigma_{\mu}(\Qvar) \otimes {\mathbb L}_{\mu}(t;\Zvar).
}
\end{equation}
Using Hopf chains, and the expression computed by Fran\c{c}ois for $n\le 6$ we arrive at the conclusion that $\sigma_{\mu}(\Qvar)$ is not Schur positive but it should be monomial positive.
In fact, for any $\mu\subseteq\delta_n$, we conjecture that it should be possible to find a subset $G_\mu$ of good chain among the Hopf chain with top element $\mu$, and a multistatistic composition ${\mathrm{mcol}}({\bf c})\models n$ for any ${\bf c}\in G_\mu$ such that
\begin{equation}\label{eq:chainstat}
\blue{
\sigma_\mu(\Qvar) = \sum_{{\bf c} \in G_\mu} M_{{\mathrm{mcol}}({\bf c})}(\Qvar),
}
\end{equation}
where $M_\alpha$ are the monomial quasisymmetric functions.
\end{conjecture}

Using some reasonable additional heuristic, we deduce the following properties of the $\sigma_\mu(\Qvar)$.
For a symmetric function $f=\sum c_\nu s_\nu$, let $\ell(f) = \max \set{\ell(\nu)}{c_\nu \ne 0}$ and $\deg(f) = \max\set{|\nu|}{c_\nu \ne 0}$.
The area of a $\mu\subseteq\delta_n$ is $\area(\mu) = {n \choose 2}-|\mu|$.

\begin{property}\label{P:len_deg}
\blue{
$\ell(\sigma_\mu) = n-|\set{k}{k \in \mu}|$} and \blue{$\deg(\sigma_\mu)\le \area(\mu)$.
}
\end{property}

Here, $|\set{k}{k\in\mu}|$ is the number of distinct parts in $\mu$ including $0$'s.
This property follows from the definition of the module of multivariate diagonal harmonics.
We also derive from the module that

\begin{property} \label{P:alt}
\blue{$\sigma_{\delta_n}=1$} and \blue{$\sigma_{0^n} = \dotprod{{\mathcal E}_n(\Qvar;\Zvar)}{s_{1^k} (\Zvar)} = {\mathcal A}_n(\Qvar)$} the $\GL$-character of the space of alternating multidiagonal harmonics.
\end{property}

We decided {\bf not to use} the property $\sigma_\mu = \sigma_{\mu'}$, since we cannot really justify it. 

\medskip
The next property is inspired from the behaviour of (Hopf) chains in the Tamari lattice. For $\mu\subseteq\delta_n$, suppose that $\mu_i=n-i$, that is the Dyck path return to the diagonal, then we can write $\mu=\alpha \odot \beta= [(n-i)^{n-i}+\beta]\cdot\alpha$, that is we decompose the dyck path of $\mu$ as the concatenation of the path of $\alpha$ followed by the path of $\beta$.

\begin{property}\label{P:multi}
If $\mu=\alpha \odot \beta$, then \blue{$\sigma_\mu=\sigma_\alpha\sigma_\beta$}.
\end{property}

We can take the scalar product of ${\mathcal E}_n(\Qvar;\Zvar)$ with different symmetric functions, this gives us identities among the $\sigma_\mu$.
It is enough to take the scalar product with Schur function (as it is a basis and we get all possible information.
For this we use the ${\mathcal E}_n(\Qvar;\Zvar)$ computed with the program $n\le 6$ in the left-hand side and use Equation~\eqref{eq:multishuff}. 

\begin{property}\label{P:coeff}
For any $\lambda \vdash n$,
\[
\blue{
\dotprod{{\mathcal E}_n(\Qvar+t;\Zvar)}{s_\lambda(\Zvar)} = \sum_{\mu\subseteq \delta_n} \dotprod{{\mathbb L}_\mu(t;\Zvar)}{s_\lambda(\Zvar)} \, \sigma_\mu(\Qvar).
}
\]
\end{property}

Notice that this gives us a property for each $\lambda\vdash n$. Our old property~{\bf 9} can be derived using~$s_{1^n}$. We have 
\begin{equation}\label{eq:alt}
\blue{
\sigma_{0^n}(\Qvar+t) =  \dotprod{{\mathcal E}_n(\Qvar+t;\Zvar)}{s_{1^n} (\Zvar)} = \sum_{\mu\subseteq \delta_n} t^{\dinv(\mu)} \sigma_\mu(\Qvar).
}
\end{equation}
This follows from the fact that $\dotprod{{\mathbb L}_\mu(t;\Zvar)}{s_\lambda(\Zvar)} = t^{\dinv(\mu)}$, since there is a unique parking function $\pi$ of shape $(\mu+1^n)/\mu$ that will give us $\comp(\pi)=1^n$ and this is the coefficient of $F_{1^n}=s_{1^n}$ which  appear only there. The $\dinv(\pi)$ of that special parking function is what we define to be $\dinv(\mu)$.

\subsection{On $\Delta$ conjecture and Tamari localization}

Now we will consider a local version of the $\Delta$-conjecture. Haglund stated that
\begin{equation}\label{eq:hag}
(\Delta'_{e_k} e_n)(q,t;\Zvar) = \sum_{\mu\subseteq\delta_n} \Big(\sum_{J\subseteq [n-1] \atop |J|=k, \, \desc(\mu)\subseteq J} q^{\sum_{i\in J} a(\mu,i)}\Big) {\mathbb L}_\mu(t;\Zvar)
\end{equation}
where $a(\mu,i)=n-i-\mu_i$ is the contribution to the area in row $i$, and~$\desc(\mu) \eqdef \set{i \in [n-1]}{\mu_i > \mu_{i+1}}$.
On the other hand, Fran\c{c}ois conjectured that~$e_{n-1-k}^{\perp_1}{\mathcal E}_n(\Qvar;\Zvar)$ should be a generalization of the $\Delta$ conjecture [\green{We have more on this later as discovered by Vincent}]. More precisely
\begin{equation}\label{eq:hag}
\Delta'_{e_k} e_n = (e_{n-1-k}^{\perp_1}{\mathcal E}_n)(q+t;\Zvar)
\end{equation}

If we expand the right-hand side using Equation~\eqref{eq:multishuff}, and use \eqref{eq:hag} for the left-hand side we get
\begin{equation}\label{eq:local}
\sum_{\mu\subseteq\delta_n} \Big(\sum_{J\subseteq [n-1] \atop |J|=k, \, \desc(\mu)\subseteq J} q^{\sum_{i\in J} a(\mu,i)}\Big) {\mathbb L}_\mu(t;\Zvar)
= \sum_{\mu\subseteq\delta_n} (e_{n-1-k}^{\perp}\sigma_{\mu})(q){\mathbb L}_{\mu}(t;\Zvar)
\end{equation}
\green{WARNING} The ${\mathbb L}_{\mu}(t;\Zvar)$ are not linearly independent but we will assume that their coefficient are equal.
Nantel justify this as the equations must be shadows of noncommutative version of LLT polynomials defined in the Catalan Hopf algebra related to Tamari lattice.
We will do the same later with non-commutative symmetric functions related to the (dual) boolean lattice.

Taking the coefficient of ${\mathbb L}_{\mu}(t;\Zvar)$ on both side of Equation~\eqref{eq:local} give what we call the Tamari localization of the $\Delta$-conjecture:

\begin{property}\label{P:localDelta}
For $\mu\subseteq\delta_n$
\[
\blue{(e_{n-1-k}^{\perp}\sigma_{\mu})(q) = \sum_{J\subseteq [n-1] \atop |J|=k, \, \desc(\mu)\subseteq J} q^{\sum_{i\in J} a(\mu,i)} 
}
\]
\end{property}

\subsection{Projection of Tamari on the (dual) boolean lattice}

For the next properties, we derive a very surprising identity from Equation~\eqref{eq:multishuff}. 
Before we need to set a few more notations.
Given $\mu \subseteq \delta_n$ we  get its descent set~$\desc(\mu) \subseteq [n-1]$.
It is well known that for a fix $A\subseteq [n-1]$, the set $\set{\mu}{\desc(\mu) = A} = [\mu_A^0, \mu_A]$ is an interval in the Tamari lattice:
\begin{itemize}
\item $\mu_A^0$ denotes the bottom of the interval, it is the bounce path defined by $A$,
\item $\mu_A$ denotes the top of the interval, it is the steep path defined by $A$.
\end{itemize}
The map $\desc$ is in fact a lattice projection from the Tamari lattice to the boolean lattice.
We sometime use composition to describe $A\subseteq[n-1]$, this has the advantage that~$n$ is understood from the composition whereas~$A$ alone does not know~$n$.
We will denote by~$c(\mu)$ the composition subset~$\desc(\mu)\subseteq [n-1]$.

Let $r_{\alpha}(\Zvar)$ be the ribbon-schur symmetric function defined by the composition $\alpha\models n$. 
For example, if $n = 6$ and $A = \{2,4,5\}$, then $\alpha = 2211$ is the composition associated to $A$.
The ribbon-Schur $r_\alpha = s_{3332/221}$ is a skew Shur function, the shape is a connected ribbon shape with $\alpha_i$ cell in row $i$.

When we let $t=0$, the polynomial LLT behave very nicely.

\begin{proposition}
\begin{equation}\label{eq:LLT0}
\blue{
{\mathbb L}_\mu(0;\Zvar) = 
\begin{cases} 0 & \text{\rm if $\mu<\mu_{\desc(\mu)}$,} \\ r_{c(\mu)}(\Zvar) & \text{\rm otherwise}.\end{cases}
}
\end{equation}
where~$\Pi_\mu$ is the set of parking function of shape~$(\mu+1^n)/u$ (standard tableau).
\end{proposition}      

\begin{proof}
Recall that the definition of LLT is
\[
{\mathbb L}_\mu(t;\Zvar) = \sum_{\pi\in \Pi_\mu} t^{\dinv(\pi)} F_{\comp(\pi)}(\Zvar)
\]
where $\dinv(\pi)$ is XXXX(hahaha), $\comp(\pi)$ is the composition associated with the descent set of the inverse of the diagonal reading the the entry of $\pi$ \red{[WARNING: do not confuse this with $c(\mu)$]} and $F$ are the Gessel fundamental quasisymmetric functions.
Now, if we compute $\dinv(\pi)$ we realize that $\dinv(\pi)>0$ always if $\mu<\mu_{\desc(\mu)}$.
That is a consequence to the fact that if we have two columns that are not consecutive than there must be at least one inversion in $\dinv(\pi)$. 

Now if we have $\mu_A$ for some $A$, then the columns of $(\mu+1^n)/u$ are consecutive. 
The parking functions of shape $(\mu+1^n)/u$ that have no dinv are exactly the filling of $(\mu+1^n)/u$ that fit the associated ribbon shape.
That shows the second part of the proposition.
\end{proof}

We can now evaluate both sides of Equation~\eqref{eq:multishuff} with $t=0$ and obtain a very surprising identity:

\begin{conjecture}\label{conj:cube}
We have an equation of ${\mathcal E}$ on the (dual) boolean lattice, namely: 
\begin{equation}\label{eq:cube}
\blue{
{\mathcal E}_n(\Qvar;\Zvar) = \sum_{A\subseteq [n-1] } \sigma_{\mu_A}(\Qvar)\otimes r_{c(\mu_A)}(\Zvar).
}
\end{equation}
\end{conjecture}

This is a totally new expression and will have many deep consequence. Francois conjecture something a bit crazy about the coefficient
above, but this would be fantastic. In particular it would clarify what we want as a generalization of the zeta map and its role with our statistics.


\begin{conjecture}\label{conj:coproduct}
There exists a map $\zeta_A$, a generalization of  the zeta map when (exactly the zeta map when to $A=\emptyset$), such that
\begin{equation}\label{eq:cube} 
\blue{
\Delta  \sigma_{\mu_A} = \sum_{\nu\le \mu_A} \sigma_\nu \otimes \sigma_{\zeta_A(\mu)}
}
\end{equation}
\end{conjecture}


The conjectures above should be the shadow on the (dual) boolean lattice of the expression on the Tamari lattice, via a Galois correspondence.
This conjecture is a consequence of Conjecture~\ref{conj:master}. We  now put $t=1$ in Equation~\eqref{eq:multishuff}  and $\Qvar+1$ in \eqref{eq:cube}.
We have to remark that LLT symmetric function at $t=1$ is a product of elementary symmetric function, but we want to expand it as a positive sum of ribbon 
Schur function (\green{WARNING: this is not unique, we do it in a non-commutative way}). Let $B = \desc(\nu)$. We have
\[
{\mathbb L}_\nu(1;\Zvar)= e_{\alpha}=\sum_{A\subset B} r_{c(A)}(\Zvar).
\]
The second equality is a direct consequence of jeux de taquin. Let $\alpha=c(A)$, let $t=1$ in Equation~\eqref{eq:multishuff} and take the coefficient of $r_\alpha$
(\green{WARNING: the $r_\beta$ are not linearly independent but we imagine that this in written in non-commutative symmetric functions.}). We compare this to the coefficient of $r_\alpha$ in Equation~\eqref{eq:cube} with $\Qvar+1$. We get
\[
\sigma_{\mu_A}(\Qvar+1) = \sum_{\nu: A\subseteq \desc(\nu)} \sigma_\nu(\Qvar)= \sum_{\nu\le \mu_A} \sigma_\nu(\Qvar).
\]
A (dual) boolean lattice localization (coefficient of $r_{c(A)}$) at $\Qvar+1$   gives our old Property~{\bf 9$'$}:

\begin{property}\label{P:steepQ1}
For any $A\subseteq [n-1]$, let $\mu_A$ be the steep shape corresponding to $A$:
\[
\blue{\sigma_{\mu_A}(\Qvar+1) = \sum_{\nu\le \mu_A} \sigma_\nu(\Qvar)
}
\]
\end{property}

\subsection{Revisiting $\Delta$ conjecture with the (dual) boolean lattice equation}

Another use of \eqref{eq:cube} is via the following identity that Fran\c{c}ois proved:
\[
e_k^\perp \sigma_{0^n}(\Qvar) = \dotprod{{\mathcal E}_n(\Qvar;\Zvar)}{s_{(k | n-k-1)}(\Zvar)},
\]
using Frobenius notation $(k | n-k-1)$ for the hook shape partition $(k+1)1^{n-k-1}$.
We remark that
\[
\dotprod{r_{c(A)}}{s_{(k | n-k-1)}} = \delta_{|A|=k},
\]
hence taking the scalar product $\dotprod{{\mathcal E}_n(\Qvar;\Zvar)}{s_{(k | n-k-1)}(\Zvar)}$ using \eqref{eq:cube} gives

\begin{property}\label{P:DeltaAlt}
For $k\le n-1$
\[
\blue{e_k^\perp \sigma_{0^n}(\Qvar)  = \sum_{A:\  |A|=k} \sigma_{\mu_A}
}
\]
\end{property}

This is a reflection of what should be a complete general $\Delta$-conjecture.
From Fran\c{c}ois expression above and the Equation~\eqref{eq:hag}, we say that $e_{k}^{\perp_1}{\mathcal E}_n(\Qvar;\Zvar)$ should be expressible as a generalization of the $\Delta$-conjecture.
Before we do so, let us use 
now the Equation~\eqref{eq:cube} in Equation~\eqref{eq:hag}:
\begin{equation}\label{eq:deltacube}
\Delta'_{e_k} e_n = (e_{n-1-k}^{\perp_1}{\mathcal E}_n)(q,t;\Zvar)
= \sum_{A\subseteq [n-1]} ( e_{n-1-k}^\perp \sigma_{\mu_A} ) (q,t) \, r_{c(\mu_A)}
\end{equation}

\begin{property}\label{P:DeltaCube}
For $k\le n-1$
\[
\blue{
\sum_{A\subseteq [n-1]} \Big( e_{n-1-k}^\perp \sigma_{\mu_A} \Big) (q,t) \, r_{c(\mu_A)}(\Zvar)  = 
\sum_{\mu\subseteq\delta_n} \Big(\sum_{J\subseteq [n-1] \atop |J|=k, \,  \desc(\mu)\subseteq J} q^{\sum_{i\in J} a(\mu,i)}\Big) {\mathbb L}_\mu(t;\Zvar)
}
\]
\end{property}

The Property~\ref{P:localDelta} is in spirit a Tamari localization of $t=0$ in the above expression. We can also set $t=1$ and do a (dual) boolean lattice localization at $A\subseteq[n-1]$:
\begin{equation}\label{eq:cube_Delta_qt}
\blue{
\Big( e_{n-1-k}^\perp \sigma_{\mu_A} \Big) (q,1) = 
\sum_{\nu\le \mu_A} \Big(\sum_{J\subseteq [n-1] \atop |J|=k, \, \desc(\nu)\subseteq J} q^{\sum_{i\in J} a(\mu,i)}\Big)
}
\end{equation}

\subsection{Anklet and generalized $\Delta$-conjecture}

The anklet~$\ank({\bf c})$ of a Hopf chain~${\bf c} = (c_1, \dots, c_r)$ is the maximal number of paths that can be inserted between the first two paths~$c_1$ and~$c_2$ of the chain~${\bf c}$ so that it remains a Hopf chain.
Note that the anklet looks like the area of the second path~$c_2$, but it is not since the rest of the chain matters.

The anklet statistic can be decomposed according to the level of the flips.
Namely, given a Hopf chain~${\bf c} = (c_1, \dots, c_r)$, consider a Hopf chain~${\bf C}$ containing~$c$ together with $\ank({\bf c})$ additional paths in between the first two paths~$c_1$ and~$c_2$ of~${\bf c}$.
Let $\ank_p({\bf c})$ be the number of flips at the horizontal level~$p$ that were made in the part of~${\bf C}$ between the first two paths of~${\bf c}$.
This decomposes the anklet into $\ank({\bf c}) = \ank_1({\bf c}) + ... + \ank_n({\bf c})$.
(\green{WARNING: it is not clear that this decomposition is independent of the choice of the superchain~${\bf C}$.})
It yields the following generalized $\Delta$-conjecture:
\[
\blue{
(e_{n-1-k}^\perp {\mathcal E}_n)(q, t, 1^{r-2};\Zvar) = \sum_{{\bf c} \text{ good chain} \atop \text{of length } r} \sum_{J \subseteq [n-1] \atop |J| = k, \, \desc({\bf c}) \subseteq J} q^{\sum_{p \in J} \ank_p({\bf c})} \otimes {\mathbb L}_{c_r}(t;\Zvar),
}
\]
where~$\desc({\bf c}) = \set{i}{\ank_i({\bf c}) < \ank_{i-1}({\bf c})}$.
Note that this can also be computed using strict chains as follows:
\[
(e_k^\perp {\mathcal E}_n)(q, t, 1^{r-2};\Zvar) = \sum_\ell \sum_{{\bf c} \text{ strict good} \atop \text{chain of length } \ell} \sum_{J \subseteq [n-1] \atop |J| = k, \, \desc({\bf c}) \subseteq J} \!\!\!\! \bigg( \one_{k = 0} \binom{r-2}{\ell-1} + q^{\sum_{p \in J} \ank_p({\bf c})} \binom{r-2}{\ell-2} \!\! \bigg) \otimes {\mathbb L}_{c_r}(t;\Zvar).
\]
This was checked by computer for~$k < n \le 4$. For $n = 5$, we do not have the right notion of good chains.

\subsection{Some new combinatorial identities}

When we evaluate Equation~\eqref{eq:cube} at $\Qvar\mapsto q$ or $\Qvar\mapsto q+t$ we get some new combinatorial identities that one should be able to prove independently of our conjecture.
For $\Qvar\mapsto q$, the symmetric function ${\mathcal E}_n(q;\Zvar)$ is the graded action of $S_n$ on the cohomology of the Flag manifold, or the graded action of $S_n$ on the quotient of polynomials in one set of variables by the ideals of symmetric function.
This is given by the Hall-Littlewood symmetric function indexed by $1^n$.
Now evaluate Equation~\eqref{eq:cube} at $\Qvar\mapsto q$, you get
\begin{equation}\label{eq:onevar}
\blue{
\frac{h_n\big[ {\Zvar}/{(1-q)}\big]}{h_n\Big[ {1}/{(1-q)}\Big]} = \sum_{A\subseteq[n-1]} q^{\area(\mu_A)} r_{c(\mu_A)}(\Zvar).
}
\end{equation}
Is this expression known? Nantel has done this expression before (unpublished?) and think Lascoux has done it also a long time ago. 
This follow from  the shuffle theorem in $q,t$, putting $t=0$. Or equivalently, put $k=n-1$ and $t=0$ in Formula~\eqref{eq:hag}.

Now at $\Qvar\mapsto q+t$ this is {\bf really} new. Evaluating Equation~\eqref{eq:cube} at $\Qvar\mapsto q$ gives us
\begin{equation}\label{eq:onevar}
\blue{
\nabla e_n= \sum_{A\subseteq[n-1]} \sigma_{\mu_A}(q,t) r_{c(\mu_A)}.
}
\end{equation}
But even more combinatorially, if we use Equation~\eqref{eq:chainstat} evaluated at $q+t$ we get at most the two chains:
\begin{equation}\label{eq:onevar}
\sigma_{\mu_A}(q,t) = \sum_{\nu\le \mu_A} q^{\ank(\delta_n,\nu,\mu_A)} t^{\beta(\delta_n,\nu,\mu_A)}.
\end{equation}
The statistic $\ank(\delta_n,\nu,\mu_A)$ is the anklet statistic. It is almost the area of $\nu$? At least when we set $t=1$ and $k=n-1$ in Property~\ref{P:DeltaCube} we get
\[
\sigma_{\mu_A}(q,1)=\sum_{\nu\le\mu_A} q^{\area(\nu)},
\]
and this agrees with Property~\ref{P:steepQ1}. Hence we conjecture
\begin{conjecture}\label{conj:area}
Given $A\subseteq[n-1]$ and $\nu\le \mu_A$
\[
\ank(\delta_n,\nu,\mu_A)=\area(\nu).
\]
\end{conjecture}

\begin{proof}[Idea of proof]
There is a unique maximal Hopf chain $(\delta_n,c_1,\ldots,c_{\ell-1},\nu)$ such that $\ell=\area(\nu)$.
We know that chain well.
We just need to show that $(\delta_n,c_1,\ldots,c_{\ell-1},\nu,\mu_A)$ is Hopf!?

\green{SECOND ARGUMENT}: If we use Conjecture~\ref{conj:coproduct} something nice happen. 
\[
\sigma_{\mu_A}(q+t) = \sum_{\nu\le \mu_A} \sigma_\nu(q) \sigma_{\zeta_A(\nu)}(t) = \sum_{\nu\le \mu_A} q^{\area(\nu)} t^{\area(\zeta_A(\nu))}.
\]
This not only show Conjecture~\ref{conj:area}, \green{but also say that $\beta(\delta_n,\nu,\mu_A)=\area(\zeta_A(\nu))$}
\end{proof}

Regardless of Conjecture~\ref{conj:area} being true or false, we should work hard to understand the values of $\ank$ and $\beta$ for this case and show 
\begin{equation}\label{eq:onevar}
\blue{
\nabla e_n= \sum_{A\subseteq[n-1]} \Big[\sum_{\nu\le \mu_A} q^{\ank(\delta_n,\nu,\mu_A)} t^{\beta(\delta_n,\nu,\mu_A)}\Big] r_{c(\mu_A)}.
}
\end{equation}
It would give us so much insight on the statistic we are seeking and $\nabla e_n$ is well studied. We should look at $\sigma_{\mu_A}(1,t)$...
put
\[
\sum_{A\subseteq[n-1]} \sigma_{\mu_A}(1,t) r_{c(\mu_A)}(\Zvar) = \sum_{\mu} {\mathbb L}_\mu(t;\Zvar) = (\nabla e_n)(1,t;\Zvar).
\]
The expression on the right is well known (but Nantel forgot).

\subsection{Some open questions and dreaming bigger}

\begin{enumerate}
\item Can we do localization for other lattices in between Tamari and the (dual) boolean lattice? [Nantel expec this to work if there is an associated Hopf algebra]. We can push the idea in the other direction, Is there an analogue of Formula~\eqref{eq:multishuff} over permutations? We can then localize over the Bruhat order? That would give us even more refined properties
\item Our exploration suggests that we should be able to construct the $\sigma_\mu$ via operators applied to $\bf 1$. These operators could be ``dual" to the Carlson-Mellit operator?
\item We feel there should be a collar version of dinv
\item We have to understand better what happens when the variable move from one side of the tensor to the other side in the expression ${\mathcal E}_n(\Qvar;\Zvar)$. Fran\c{c}ois' experiments suggest (explicit expressions for $n\leq 4$, and partial for $n=5$) that there is in fact a nice expansion of the form
\begin{equation}
    \mathcal{E}_n(\bm{q}+\bm{t};\bm{z}) = \sum_{\mu\subseteq \delta_n} \sigma_\mu(\bm{q}) \mathbb{L}_\mu(\bm{t};\bm{z}),
  \end{equation}
  with $\bm{t}$ a general set of variables. To get these it is interesting to exploit the specialization that corresponds to  Conjecture~\ref{conj:master} (all left terms of length $1$ in $\mathbb{L}_\mu(\bm{t};\bm{z})$ correspond to classical LLT: $\mathbb{L}_\mu(t;\bm{z})$), as well as the fact that 
  the formula has to be $(\bm{q},\bm{t})$-symmetric, so that one has 
     $$\sum_{\mu\subseteq \delta_n} \sigma_\mu(\bm{q}) \mathbb{L}_\mu(\bm{t};\bm{z})=\sum_{\mu\subseteq \delta_n} \sigma_\mu(\bm{t}) \mathbb{L}_\mu(\bm{q};\bm{z}).$$
     Calculation suggest that there are nice stabilities (to be described later) in the ``coefficients''.
\end{enumerate}

%-------------------------------------------------------------------
% Begin BIBLIOGRAPHY
%-------------------------------------------------------------------

%\small
%\bibliographystyle{abbrv}  
%%\bibliographystyle{amsalpha}  
%%\bibliographystyle{plain}  
%%\bibliographystyle{amsplain}  
%\bibliography{}

 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\auteur#1{{\sc #1}}
\def\titreref#1{{\em #1}}
\def\vol#1{{\bf #1}}

\renewcommand{\refname}{{References}}
\begin{thebibliography}{10}  

\bibitem{ALW}
\auteur{D.\,Armstrong, N.\,Loehr, and G.\,Warrington},
\titreref{Rational parking functions and Catalan numbers},
Annals of Combinatorics, 20 (2016), no. 1, 21--58. 
See \href{http://arxiv.org/abs/1403.1845}{arXiv:1403.1845}. \MR{3461934}. 
 
\bibitem{reiner}
\auteur{D.\,Armstrong, V.\,Reiner, and B.~Rhoades},
\titreref{Parking spaces},
Advances in Mathematics, 269 (2015), 647--706. 
See \href{http://arxiv.org/abs/1204.1760}{arXiv:1204.1760}. \MR{3281144}.

\bibitem{armstrong}
\auteur{D.\,Armstrong, B.\,Rhoades, and N.\,Williams},
\titreref{Rational associahedra and noncrossing partitions},
Electronic Journal of Combinatorics,  20 (2013), no. 3, Paper 54, 27 pp. \\
   See \href{http://arxiv.org/abs/1305.7286}{arXiv:1305.7286}. \MR{3118962}.

 \bibitem{open}
\auteur{F.\,Bergeron},
\titreref{Open Questions for Operators Related to Rectangular Catalan Combinatorics},  
Journal of Combinatorics  Vol. 8, No. 4 (2017),   673--703. 
See \href{http://arxiv.org/abs/1603.04476}{arXiv:1603.04476}. \MR{3682396}.


\bibitem{multicomplex}
\auteur{F.\,Bergeron},
\titreref{Multivariate Diagonal Coinvariant Spaces for Complex Reflection Groups}, 
Advances in Mathematics, 239 (2013), 97--108. 
See \href{http://arxiv.org/abs/1105.4358}{arXiv:1105.4358}. \MR{3045143}.

\bibitem{bergeron}
\auteur{F.\,Bergeron},
\titreref{Algebraic Combinatorics and Coinvariant Spaces}, 
CMS Treatise in Mathematics, CMS and A.K.Peters,  2009.

\bibitem{ScienceFiction}
\auteur{F.\,Bergeron and A.\,M. Garsia},
\titreref{Science Fiction and Macdonald's Polynomials}, 
in Algebraic Methods and $q$-Special Functions, J.P. Van Dejen, L.Vinet (eds.), 
CRM Proceedings \& Lecture Notes, AMS (1999), 1--52. \MR{1726826}.

\bibitem{identity}
\auteur{F.\,Bergeron, A.\,M. Garsia, M.\,Haiman, and G.\,Tesler},
\titreref{Identities and Positivity Conjectures for some remarkable Operators in the Theory of Symmetric Functions},
Methods and Applications of Analysis, 
 Volume 6 Number 3 (1999), 363--420. \MR{1803316}.

\bibitem{compositionalshuffle}
\auteur{F.\,Bergeron, A.\,M. Garsia, E.\,Leven,  and G.\,Xin},
\titreref{Compositional \lowercase{$(km,kn)$}--Shuffle Conjectures},
International Mathematics Research Notices, (2015), 33 pages.
\href{https://doi.org/10.1093/imrn/rnv272}{doi.org/10.1093/imrn/rnv272}. See also:
\href{http://arxiv.org/abs/1404.4616}{arXiv:1404.4616}.  \MR{3556418}.


\bibitem{trivariate} 
\auteur{F.\,Bergeron, and L.-F.\,Pr\'eville-Ratelle},
\titreref{Higher Trivariate Diagonal Harmonics via generalized Tamari Posets}, 
Journal of Combinatorics, \vol{3} (2012), 317--341. 
See \href{http://arxiv.org/abs/arXiv:1105.3738}{arXiv:1105.3738}. \MR{3029440}.

\bibitem{hopf} 
\auteur{N.\,Bergeron, C.~Ceballos, and V.~Pilaud},
\titreref{Hopf Dreams}, 
See \href{https://arxiv.org/abs/1807.03044}{https://arxiv.org/abs/1807.03044}

\bibitem{bizley}
\auteur{T. L. Bizley},
\titreref{Derivation of a new formula for the number of minimal lattice paths from $(0,0)$ to $(km,kn)$ having just $t$ contacts with the line and a proof of Grossman’s formula for the number of paths which may touch but do not rise above this line}, J. Inst. Actuar. 80, (1954), 55--62.


\bibitem{geramita}
 \auteur{M.\,Boij and A.\,Geramita},
\titreref{Notes on Diagonal Coinvariants of the Dihedral Group},
Canadian Mathematical Bulletin,
\href{https://doi.org/10.4153/CMB-2010-066-7}{doi.org/10.4153/CMB-2010-066-7}. 

\bibitem{BurbanSchiffmann} 
\auteur{I. Burban and O. Schiffmann},
\titreref{On the Hall algebra of an elliptic curve I},
Duke Mathematical Journal,  \vol{161} (2012), 1171--1231. \MR{2922373}.

\bibitem{carlsson_mellit}
\auteur{E.\,Carlsson, and A.\,Mellit},
\titreref{A Proof of the Shuffle Conjecture},
Journal of the American Mathematical Society, Published electronically: November 30, 2017.\\
\href{https://doi.org/10.1090/jams/893}{doi.org/10.1090/jams/893}.
See also: \href{http://arxiv.org/abs/arXiv:1508.06239}{arXiv:1508.06239}.

\bibitem{cherednik}
\auteur{I.\,Cherednik},
\titreref{Jones polynomials of torus knots via DAHA},
International Mathematics Research Notices, Volume 2013, Issue 23, 1 January 2013, Pages 5366--5425.
See \href{https://arxiv.org/abs/1111.6195}{arXiv:1111.6195}.

\bibitem{desarmenien}
\auteur{J.\,D\'esarm\'enien, B.\,Leclerc and J.-Y.\,Thibon},
\titreref{Hall-Littlewood Functions and Kostka-Foulkes Polynomials in Representation Theory},
\href{http://www.mat.univie.ac.at/~slc/}{S\'eminaire Lotharingien de Combinatoire} 32 (1994), 
Art. B32c, 38 pp. \MR{1399504}.

\bibitem{dunfield}
\auteur{N.M.\,Dunfield, S.\,Gukov, and J.\,Rasmussen},
\titreref{The Superpolynomial for Knot Homologies},
Experimental Mathematics, 15 (2006) 129--159.
See \href{https://arxiv.org/abs/math/0505662}{arXiv:0505662}. \MR{2253002}.

\bibitem{farb}
\auteur{T.\,Church, J.S.\,Ellenberg, and B.\,Farb}
\titreref{FI-modules and stability for representations of symmetric groups},
Duke Mathematical Journal, 
Volume 164, Number 9 (2015), 1833--1910.
See \href{https://arxiv.org/pdf/1204.4533.pdf}{arXiv:1204.4533}. \MR{3357185}.

\bibitem{leven}
\auteur{A.M.\,Garsia, E.\,Leven, N.\,Wallach, and G.\,Xin},
\titreref{A new Plethystic Symmetric Function Operator
and The rational Compositional Shuffle Conjecture at $t=1/q$},
Journal of Combinatorial Theory, Series A, 145 (2017), 57--100.
\href{https://arxiv.org/pdf/1501.00631.pdf}{arXiv:1501.00631}. \MR{3551646}.

\bibitem{garsia_haiman} 
  \auteur{A.M.\,Garsia and M.\,Haiman},
 \titreref{A Remarkable $q,t$-{C}atalan Sequence and $q$-Lagrange Inversion}, 
 Journal of Algebraic Combinatorics, \vol{5} (1996), no.\,3, 191--244. \MR{1394305}.

  \bibitem{GorskyNegut}
\auteur{E.\,Gorsky and A.\,Negut},
\titreref{Refined Knot Invariants and Hilbert Schemes},
Journal de Math\'ematiques Pures et Appliqu\'ees,
Volume 104, Issue 3 (2015),  403--435.
See \href{http://arxiv.org/abs/1304.3328}{arXiv:1304.3328}.  \MR{3383172}.

\bibitem{haglund_knot}
\auteur{J.\,Haglund},
\titreref{\href{https://www.math.upenn.edu/~jhaglund/preprints/IMAfinal.pdf}{The combinatorics of knot invariants arising from the study of Macdonald polynomials}}. 
Recent trends in combinatorics, 579--600, IMA Vol. Math. Appl., 159, Springer, 2016. \MR{3526424}.

%\bibitem{HHLRU}
%\auteur{J.\,Haglund, M.\,Haiman, N.\,Loehr, J.\,Remmel, and A.\,Ulyanov},
%\titreref{A Combinatorial Formula for the Character of the Diagonal Coinvariants},
%Duke Mathematical Journal,  126 (2005), 195--232. \MR{2115257}.

\bibitem{haglund}
\auteur{J.\,Haglund,  J.\,Remmel, and A.\,Wilson}, 
\titreref{The Delta Conjecture},  
Transactions of the American Mathe\-matical Society, (accepted).
\href{https://doi.org/10.1090/tran/7096}{doi.org/10.1090/tran/7096} 
See \href{http://arxiv.org/abs/1509.07058}{arXiv:1509.07058}.

\bibitem{HHLRU}
\auteur{J.\,Haglund, M.\,Haiman, N.\,Loehr, J.\,Remmel, and A.\,Ulyanov},
\titreref{A Combinatorial Formula for the Character of the Diagonal Coinvariants},
Duke Mathematical Journal,  126 (2005), 195--232. \MR{2115257}.


\bibitem{haimanhilb}    
\auteur{M.\,Haiman},
\titreref{Combinatorics, symmetric functions and Hilbert schemes},  
In CDM 2002: Current Developments in Mathematics in Honor of  Wilfried Schmid \& George Lusztig, 
International Press Books  (2003) 39--112. \MR{2051783}. 

\bibitem{haimanvanishing}
\auteur{M.\,Haiman}, 
\titreref{Vanishing Theorems and Character Formulas for the Hilbert Scheme of Points in the Plane}, 
Inventiones Mathematicae 149 (2002), 371--407. \MR{1918676}.


\bibitem{hogancamp}
\auteur{M.\,Hogancamp},
\titreref{Khovanov-Rozansky Homology and Higher Catalan Sequences}, (2017).\\
See \href{http://arxiv.org/abs/arXiv:1704.01562}{arXiv:1704.01562}.

\bibitem{kirillov}
\auteur{A.N.\,Kirillov},
\titreref{New Combinatorial Formula For Modified Hall–Littlewood Polynomials}.
$q$-series from a contemporary perspective (South Hadley, MA, 1998), 283--333, 
Contemporary Mathematics, 254, Amer. Math. Soc., Providence, RI, 2000. \MR{1768934}.


\bibitem{lascoux}
\auteur{A.\,Lascoux},
\titreref{Cyclic permutations on words, tableaux and harmonic polynomials}, 
Proceedings of the Hyderabad Conference on Algebraic Groups, 1989, Manoj Prakashan, Madras (1991), 323--347. \\
\MR{1131318}.

%\bibitem{LascouxSchutzenberger}
%\auteur{A.\,Lascoux and M.P.\,Schützenberger}, 
%\titreref{Sur une conjecture de H.O. Foulkes}, 
%Comptes rendus de l'Acad\'emie des Sciences, Paris 286A (1978), 323--324. \MR{0472993}.

\bibitem{macdonald} 
\auteur{I.\,G. Macdonald},
\titreref{Symmetric functions and {H}all polynomials}, second ed., Oxford
Mathematical Monographs, The Clarendon Press Oxford University Press, New York, 1995, 
With contributions by A. Zelevinsky, Oxford Science Publications. \MR{3443860}.
  
 \bibitem{mellit_homology}  
 \auteur{A.\,Mellit},
 \titreref{Homology of Torus Knots}, (2107).
 See \href{http://arxiv.org/abs/arXiv:1704.07630}{arXiv:1704.07630}.
 
 
  \bibitem{mellit_braids}  
  \auteur{A.\,Mellit},
  \titreref{Toric Braids and $(m,n)$-Parking Functions},
See \href{http://arxiv.org/abs/arXiv:1604.07456}{arXiv:1604.07456}.

\bibitem{morton_samuelson}
\auteur{H.\,Morton, and P.\,Samuelson},
\titreref{The HOMFLYPT Skein Algebra of the Torus and the Elliptic Hall Algebra}, 	
Duke Mathematical Journal,  166, no. 5 (2017), 801--854.
See \href{http://arxiv.org/abs/1410.0859}{arXiv:1410.0859}.  \MR{3626565}.



\bibitem{NegutShuffle}
\auteur{A.\,Negut},
\titreref{The Shuffle Algebra Revisited},
International Mathematics Research Notices, (2014) (22): 6242--6275. 
\href{https://doi.org/10.1093/imrn/rnt156}{doi.org/10.1093/imrn/rnt156}. See also:
\href{http://arxiv.org/abs/1209.3349}{arXiv:1209.3349}, (2012). \MR{3283004}.

\bibitem{Qiu}
\auteur{D.\,Qiu, J.B.\,Remmel, E.\,Sergel, and G.\,Xin},
\titreref{On the Schur Positivity of $\Delta_{e_2} e_n$}, (2017).\\
See \href{https://arxiv.org/pdf/1710.03340}{arXiv:1710.03340}.

\bibitem{FPSAC_Qiu}
\auteur{D.\,Qiu, and J.B.\,Remmel},
\titreref{Schur Function Expansions and the Rational Shuffle Conjecture},
\href{http://www.mat.univie.ac.at/~slc/wpapers/FPSAC2017/83\%20Qiu\%20Remmel.pdf}{S\'eminaire Lotharingien de Combinatoire, 78B.83} (2017), 12 pp.

  
\bibitem{SchiffVassK}
\auteur{O.\,Schiffmann and E.\,Vasserot},
\titreref{The Elliptical Hall Algebra and the Equivariant $K$-theory of the Hilbert Scheme of $\mathbb{A}^2$},
Duke Mathematical Journal, \vol{162-2} (2013), 279--366. \MR{3018956}.

\bibitem{sheppard} 
\auteur{G.C.\,Shephard and J.A.\,Todd}, 
\titreref{Finite Unitary Reflection Groups}, 
Canadian Journal of Mathematics \vol{6} (1954),  274--304. \MR{0059914}.

\bibitem{stanley}
\auteur{R.P.\,Stanley},
\titreref{Parking Functions and Noncrossing Partitions},
The Electronic Journal of Combinatorics, The Wilf Festschrift (Philadelphia, PA, 1996). The Electronic Journal of Combinatorics 4 (1997), no. 2, \href{http://www.combinatorics.org/ojs/index.php/eljc/article/view/v4i2r20}{Research Paper 20}, 14 pages. \MR{1444167}

 \bibitem{stembridge}
 \auteur{J.\,Stembridge},
 \titreref{Multiplicity-Free Products of Schur Functions},
  Annals of Combinatorics (2001), 1--9. \MR{1904379}.
  
%  \bibitem{thiery}
%  \auteur{N.\,Thiery},
%  \titreref{Computing the bi-character of diagonal harmonic polynomials}.
 
 \bibitem{wallace}
 \auteur{N.~Wallace},
 
  \bibitem{wilson}
  \auteur{A.\,Wilson},
  \titreref{Torus Link Homology and the Nabla Operator},
   Journal of Combinatorial Theory, Series A 154 (2018), 129--144.  
   See \href{http://arxiv.org/abs/1606.00764}{arXiv:1606.00764}. \MR{3718063}.
  
  \bibitem{zabrocki}
  \auteur{M.\,Zabrocki},
  \titreref{Vertex operators for standard bases of the symmetric functions},
   Journal of Algebraic Combinatorics, 13 (2001), no. 1, 83--101.
   See \href{https://arxiv.org/pdf/math/9904084.pdf}{arXiv:904084}. \MR{1817706}.

\end{thebibliography}


\end{document}
